---
title: 'Final Project: NYC Airbnb Data Analysis'
author: "David Brehm"
date: '2021-03-04'
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(lubridate)
library(grid)
library(gridExtra)
library(caret)
library(car)
library(leaflet)

set.seed(123)
```

```{r, echo = FALSE}
setwd("D:/School/520/Final")  # Set working directory 
data <- read.csv('AB_NYC_2019.csv')
```


## $\underline{Data\space Cleaning}$

### Observe Missing Values and Categorical Variables
```{r}
colSums(is.na(data))

data %>% group_by(neighbourhood) %>% tally(sort=TRUE)
data %>% group_by(neighbourhood_group) %>% tally(sort=TRUE)
data %>% group_by(room_type) %>% tally(sort=TRUE)
```

### Convert Columns and Replace Missing Values

```{r}
review_date <- ymd(data$last_review)  # Convert last_review to year-month-day.
today <- ymd(today())  # Date today.
data$days_since_last_review <- sapply(1:48895, function(i) difftime(today[1], review_date[i]))  # Convert last review date to days since last review.

data$reviews_per_month[is.na(data$reviews_per_month)] <- 0  # Replace NA review_per_month with 0.
data$days_since_last_review[is.na(data$days_since_last_review)] <- 
  max(data$days_since_last_review[which(!is.na(data$days_since_last_review))]) + 1  # Replace NA days since last review with the max plus 1.


# Convert categorical variables to factor.
data$neighbourhood <- as.factor(data$neighbourhood)
data$neighbourhood_group <- as.factor(data$neighbourhood_group)
data$room_type<- as.factor(data$room_type)

fdata <- subset(data, select = -c(id, name, host_id, host_name, neighbourhood, last_review))  # Columns not being used in the model.
```

## $\underline{Distributions}$

```{r}
# Histograms
p1 <- qplot(price, data=fdata) + xlab("Price ($)") + ylab("Count") + ggtitle("Price")
p2 <- qplot(minimum_nights, data=fdata) + xlab("Minimum Nights") + ylab("Count") + ggtitle("Minimum Nights")
p3 <- qplot(number_of_reviews, data=fdata) + xlab("Number of Reviews") + ylab("Count") + ggtitle("Number of Reviews")
p4 <- qplot(reviews_per_month, data=fdata) + xlab("Reviews Per Month") + ylab("Count") + ggtitle("Reviews Per Month")
p5 <- qplot(calculated_host_listings_count, data=fdata) + xlab("Number of Host Listings") + ylab("Count") + ggtitle("Number of Host Listings")
p6 <- qplot(availability_365, data=fdata) + xlab("Days Available") + ylab("Count") + ggtitle("Days Available")
p7 <- qplot(days_since_last_review, data=fdata) + xlab("Days Since Last Review") + ylab("Count") + ggtitle("Days Since Last Review")

grid.arrange(p1,p2,p3,p4,p5,p6,p7, nrow=3)

ggplot(data, aes(x=neighbourhood_group, fill = room_type)) + geom_bar(position="dodge") +
  labs(x="Neighborhood Group",fill="Room Type", title="Listings by Neighborhood Group and Room Type")

ggplot(data, aes(x=neighbourhood_group, y=price,fill = room_type)) + 
  geom_bar(stat="summary",fun="mean",position="dodge") + 
  labs(x="Neighborhood Group",y="Average Price ($)", fill="Room Type", title="Average Price by Neighborhood Group and Room Type")
```

## $\underline{Predicting\space Price}$
```{r}
modData = fdata[c(5,1,4,6:11)]
mod <- lm(price ~ . , data = modData)
summary(mod)
```

### MSE and RMSE of the Model
```{r}
paste("MSE:", mean(mod$residuals^2), sep=" ")  # MSE
paste("RMSE:", sqrt(mean(mod$residuals^2)), sep=" ")  # RMSE
```


### Standardized Residuals
```{r}
modData$stdRes <- rstandard(mod)  # Standardized residuals
modData$studRes <- rstudent(mod)  # Studentized residuals
modData$largeResid <-  modData$stdRes > 2 | modData$stdRes < -2
sum(modData$largeResid)

```
### Leverage, Cooks Distance, and Covariance
```{r}
modData$lev <- hatvalues(mod)
modData$cooksDist <- cooks.distance(mod)
modData$covRatio <- covratio(mod)
modData1 <- modData[modData$largeResid, c("lev","cooksDist","covRatio")]
summary(modData1)
```
None of these values have a Cooks Distance greater than 1, so none of these cases have an undue influence on the model.


### Durbin Watson Test
```{r}
durbinWatsonTest(mod)
```
The Durbin Watson test returns a value between 0 and 4. A result closer to 0 indicates a positive autocorrelation while a result closer to 4 indicates a negative autocorrelation. The closer to 2 the better. The test for this model returns 1.848, which indicates minimal autocorrelation.


### Plotting Predicted vs Actual Price
```{r}
predicted <- predict(mod,modData[c(2:9)])
prices <- cbind.data.frame(predicted, modData$price)
ggplot(data = prices, aes(x=predicted,y=modData$price)) + geom_point(alpha = 0.01, colour="purple") + geom_abline(linetype=2) +
  ylim(0,1000) + xlab('Predicted Price') + ylab('Actual Price')
```



### Correlation
```{r}
cor(modData[,4:9])

```


### Recursive Feature Elimination
```{r}
sample <- sample_n(modData, 3000)
control <- rfeControl(functions=rfFuncs, method="cv", number=3, repeats=2)
result <- rfe(sample[,2:9], sample[,1], sizes = c(2:9), rfeControl = control)
result
predictors(result)
plot(result, type=c("g","o"))
```


## $\underline{Clustering\space and\space Map\space Plots}$
### Elbow method to choose a value for k
```{r}
dataCoord <- data[c('longitude','latitude')]
cluster2 <- kmeans(dataCoord,2)
cluster3 <- kmeans(dataCoord,3)
cluster4 <- kmeans(dataCoord,4)
cluster5 <- kmeans(dataCoord,5)
cluster6 <- kmeans(dataCoord,6)
cluster7 <- kmeans(dataCoord,7)
cluster8 <- kmeans(dataCoord,8)
cluster9 <- kmeans(dataCoord,9)
cluster10 <- kmeans(dataCoord,10)
cluster11 <- kmeans(dataCoord,11)
cluster12 <- kmeans(dataCoord,12)
data$k2 <- as.factor(cluster2$cluster)
data$k3 <- as.factor(cluster3$cluster)
data$k4 <- as.factor(cluster4$cluster)
data$k5 <- as.factor(cluster5$cluster)
data$k6 <- as.factor(cluster6$cluster)
data$k7 <- as.factor(cluster7$cluster)
data$k8 <- as.factor(cluster8$cluster)
data$k9 <- as.factor(cluster9$cluster)
data$k10 <- as.factor(cluster10$cluster)
data$k11 <- as.factor(cluster11$cluster)
data$k12 <- as.factor(cluster12$cluster)

k <- 2:12
wsslist <- c(cluster2$tot.withinss, cluster3$tot.withinss, cluster4$tot.withinss, cluster5$tot.withinss, cluster6$tot.withinss, 
                 cluster7$tot.withinss, cluster8$tot.withinss, cluster9$tot.withinss,cluster10$tot.withinss,
                 cluster11$tot.withinss,cluster12$tot.withinss)
plot(k, wsslist, type="b", pch=19, frame=FALSE, xlab="Number of clusters", ylab="Total Within-Clusters Sum of Squares")
```

### KMeans Clustering with K=5
```{r}
pal <- colorFactor(palette = c('red', 'green', 'blue', 'orange', 'purple'), domain = data$k5)
plot1 <- leaflet(data = data) %>% addTiles() %>% addCircleMarkers(radius = 0.5, color = ~pal(k5))
plot1
```

### Listings by Neighborhood Group
```{r}
pal <- colorFactor(palette = c('green', 'blue', 'red', 'orange', 'purple'), domain = data$neighbourhood_group)
plot2 <- leaflet(data = data) %>% addTiles() %>% addCircleMarkers(radius = 0.5, color = ~pal(neighbourhood_group))
plot2
```

### Listings by Price
```{r}
priceBin <- c(0, 100, 200, 300, 400, 500, 1000, 2500, 10000)
palGYR <- colorBin(palette = c("#A50026","#D73027","#F46D43","#FDAE61","#A6D96A","#66BD63","#1A9850","#006837"), bins=priceBin, reverse = TRUE)
plot3 <- leaflet(data = fdata) %>% addTiles() %>% addCircleMarkers(radius = 0.5, color = ~palGYR(price)) %>% addLegend("topleft",pal = palGYR, values = ~price, opacity = 1)
plot3
```


